---
title: <center> Statistical Analysis
pagetitle: Statistical Analysis
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: cosmo
---

```{r setup, include = FALSE}
library(tidyverse)
library(modelr)
library(table1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Overview

To learn more about what influences adults to obtain an influenza vaccine, we will conduct additional statistical analyses of the National Health Survey data set. We will be using the tidied data from our earlier [exploratory analysis](methods.html). 


## Univariate Analysis

To begin, we would like to take a better look at the distribution of the data and important variables. 

```{r table1, message = FALSE, warning = FALSE}
flu_adult_df = 
  read_csv("./final_project_large_data/flu_adult_df.csv") %>% 
  mutate(
    flu_shot = str_to_title(flu_shot),
    flu_shot = as.factor(flu_shot),
    race_eth = 
      case_when(
        nh_race == "hispanic"          ~ "Hispanic",
        nh_race == "NH AIAN"           ~ "Native or mixed race",
        nh_race == "NH AIAN and other" ~ "Native or mixed race",
        nh_race == "NH asian"          ~ "Asian",
        nh_race == "NH black"          ~ "Black",
        nh_race == "NH white"          ~ "White",
        nh_race == "multiple races"    ~ "Native or mixed race"
      ),
    last_dr_visit = 
      case_when(
        last_drvisit == "10+ years"            ~ "5+ years",
        last_drvisit == "within last 10 years" ~ "5+ years",
        last_drvisit == "within last 5 years"  ~ "<5 years",
        last_drvisit == "within last 3 years"  ~ "<5 years",
        last_drvisit == "within last 2 years"  ~ "<2 years",
        last_drvisit == "within past year"     ~ "Within past year",
        last_drvisit == "never"                ~ "5+ years"
      ),
    education = 
      case_when(
        education == "12th grade, no dipolma" ~ "Less than high school",
        education == "Associate degree-aca"   ~ "Associate's degree",
        education == "Associate degree-tech"  ~ "Associate's degree",
        education == "Bachelor's degree"      ~ "Bachelor's degree",
        education == "College, no degree"     ~ "Some college",
        education == "Doctoral degree"        ~ "Advanced degree",
        education == "GED"                    ~ "High school diploma/GED",
        education == "Grades 1-11"            ~ "Less than high school",
        education == "HS graduate"            ~ "High school diploma/GED",
        education == "Master's degree"        ~ "Advanced degree",
        education == "never attended"         ~ "Less than high school",
        education == "Professional school"    ~ "Advanced degree"
      ),
    health = str_to_title(health),
    health = as.factor(health),
    insurance = str_to_title(insurance),
    insurance = as.factor(insurance),
    usborn = str_to_title(usborn),
    usborn = as.factor(usborn),
    bmi = str_to_title(bmi),
    bmi = as.factor(bmi),
    race_eth = fct_infreq(race_eth),
    health = fct_relevel(health, "Excellent", "Very Good", "Good", "Fair", "Poor"),
    last_dr_visit = fct_relevel(last_dr_visit, "Within past year", "<2 years",
                               "<5 years", "5+ years"),
    education = fct_relevel(education, "Less than high school", 
                            "High school diploma/GED", "Some college",
                            "Associate's degree", "Bachelor's degree",
                            "Advanced degree"),
    usborn = fct_relevel(usborn, "Yes", "No"),
    bmi = fct_relevel(bmi, "Underweight", "Healthy Weight", "Overweight", "Obese")
  ) 

label(flu_adult_df$flu_shot) = "Flu Vaccine Received"
label(flu_adult_df$age) = "Age"
label(flu_adult_df$sex) = "Sex"
label(flu_adult_df$race_eth) = "Race/Ethnicity"
label(flu_adult_df$health) = "Health Status"
label(flu_adult_df$insurance) = "Insurance Coverage"
label(flu_adult_df$bmi) = "BMI"
label(flu_adult_df$last_dr_visit) = "Last Doctor Visit"
label(flu_adult_df$education) = "Education"
label(flu_adult_df$usborn) = "US-Born"
units(flu_adult_df$age) = "years"
units(flu_adult_df$bmi) = "kg/m^2"

table1(~ age + sex + race_eth + health + insurance + bmi +
         last_dr_visit + education + usborn | flu_shot, 
       data = flu_adult_df,
       caption = "Table 1: Distribution of demographic and health-related 
       covariates by vaccination status.")
```

## Logistic Regression Modeling

To further investigate the data, we will run a logistic regression to look into what variables significantly influence the odds that an individual will obtain the flu vaccine. We believe that health insurance coverage is an important predictor of vaccination behavior, so we will begin with that as a main effect and add other explanatory variables from there.

```{r lr_crude}
flu_adult_model = 
  flu_adult_df %>% 
  mutate(
    flu_shot = case_when(
      flu_shot == "Yes" ~ 1,
      flu_shot == "No"  ~ 0
    )
  ) 

logreg_crude = 
  glm(flu_shot ~ insurance, data = flu_adult_model, family = "binomial") 

lr_crude_results = 
  logreg_crude %>% 
  broom::tidy() %>% 
  janitor::clean_names() %>% 
  mutate(
    term = str_replace(term, "insurance", "Insurance Coverage: ")
  ) %>% 
  dplyr::select(-c(std_error, statistic))

lr_crude_results %>% knitr::kable(digits = 4)
```

We can see from our crude, main effects model that adults who are not covered by health insurance have `r filter(lr_crude_results, term == "Insurance Coverage: Not Covered") %>% dplyr::select(estimate) %>% exp() %>% round(2)` times the odds of getting a flu vaccine compared to individuals who do have health insurance, on average. Our coefficient for the insurance term is significant in this model at a significance level of $\alpha$ = 0.05.

To build out our model, we will include effects of patient sex, education level, race/ethnicity, health status, BMI category, recency of last doctor visit, and whether the patient is a native citizen or an immigrant. Our full parametric model is as follows:

```{r full}
logreg_full = 
  glm(flu_shot ~ insurance + sex + education + race_eth + health + bmi + 
                  last_dr_visit + usborn, 
                 data = flu_adult_model, family = "binomial") 

lr_full_results = 
  logreg_full %>% 
  broom::tidy() %>% 
  janitor::clean_names() %>% 
  mutate(
    term = str_replace(term, "insurance", "Insurance Coverage: "),
    term = str_replace(term, "education", "Education: "),
    term = str_replace(term, "race_eth", "Race/Ethnicity: "),
    term = str_replace(term, "health", "Health: "),
    term = str_replace(term, "bmi", "BMI: "),
    term = str_replace(term, "last_dr_visit", "Last Doctor Visit: "),
    term = str_replace(term, "usborn", "US Born: "),
    term = str_replace(term, "sex", "Sex: ")
  ) %>% 
  dplyr::select(-c(std_error, statistic))

lr_full_results %>% knitr::kable(digits = 4)
```

In our full adjusted model, we see that adults who are not covered by health insurance have `r filter(lr_full_results, term == "Insurance Coverage: Not Covered") %>% dplyr::select(estimate) %>% exp() %>% round(2)` times the odds of getting a flu vaccine compared to individuals who do have health insurance, on average. Our coefficient for the insurance term is significant in this model at a significance level of $\alpha$ = 0.05. Regarding the other variables in our model, all variables have statistically significant levels with the exception of BMI and US-born status. Therefore, we will remove those variables from the model.

```{r adj}
logreg_adj =
  glm(flu_shot ~ insurance + sex + education + race_eth + health + last_dr_visit, 
                 data = flu_adult_model, family = "binomial")

lr_adj_results = 
  logreg_adj %>% 
  broom::tidy() %>% 
  janitor::clean_names() %>% 
  mutate(
    term = str_replace(term, "insurance", "Insurance Coverage: "),
    term = str_replace(term, "education", "Education: "),
    term = str_replace(term, "race_eth", "Race/Ethnicity: "),
    term = str_replace(term, "health", "Health: "),
    term = str_replace(term, "last_dr_visit", "Last Doctor Visit: "),
  ) %>% 
  dplyr::select(-c(std_error, statistic))

lr_adj_results %>% knitr::kable(digits = 4)
```

In our parsimonious adjusted model, we see that adults who are not covered by health insurance have `r filter(lr_adj_results, term == "Insurance Coverage: Not Covered") %>% dplyr::select(estimate) %>% exp() %>% round(2)` times the odds of getting a flu vaccine compared to individuals who do have health insurance, on average. Our coefficient for the insurance term is significant in this model at a significance level of $\alpha$ = 0.05. All of the other variables in the model are significant, so we are satisfied with this model.

## Testing Model Fit

To compare the models, we will evaluate the AIC, Akaike's Information Criterion, for each model. The model that has a smaller AIC is a better fit for the data. 

```{r aic}
aic_test = 
  tibble(
    "Crude" = AIC(logreg_crude),
    "Full" = AIC(logreg_full),
    "Parsimonious" = AIC(logreg_adj)
  ) %>% 
  pivot_longer(
    everything(),
    names_to = "Model",
    values_to = "AIC"
  )

aic_test %>% knitr::kable()
```

As we can see, the full model has a smaller AIC than the crude and final model, meaning that it is a better fit for the data. However, because we are looking for the most parsimonious model rather than the most saturated model with the greatest fit for the data, we will select the third model as our final adjusted model. 

