---
title: "Storyboard Page"
output: flexdashboard::flex_dashboard
---


```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(haven)
library(plotly)

brfss_df =
  read_xpt("./final_project_large_data/LLCP2019.xpt") %>%
  janitor::clean_names()%>%
  mutate(
    state = as.factor(state),
    flushot_yn = as.factor(flushot7)
  )%>%
  mutate(
    state = recode(state,
         "1" = "Alabama",
         "2" = "Alaska",
         "4" = "Arizona",
         "5" = "Arkansas",
         "6" = "California",
         "8" = "Colorado",
         "9" = "Connecticut",
         "10" = "Delaware",
         "11" = "District of Columbia",
         "12" = "Florida",
         "13" = "Georgia",
         "15" = "Hawaii",
         "16" = "Idaho",
         "17" = "Illinois",
         "18" = "Indiana",
         "19" = "Iowa",
         "20" = "Kansas",
         "21" = "Kentucky",
         "22" = "Louisiana",
         "23" = "Maine",
         "24" = "Maryland",
         "25" = "Massachusetts",
         "26" = "Michigan",
         "27" = "Minnesota",
         "28" = "Mississippi",
         "29" = "Missouri",
         "30" = "Montana",
         "31" = "Nebraska",
         "32" = "Nevada",
         "33" = "New Hampshire",
         "35" = "New Mexico",
         "36" = "New York",
         "37" = "North Carolina",
         "38" = "North Dakota",
         "39" = "Ohio",
         "40" = "Oklahoma",
         "41" = "Oregon",
         "42" = "Pennsylvania",
         "44" = "Rhode Island",
         "45" = "South Carolina",
         "46" = "South Dakota",
         "47" = "Tennessee",
         "48" = "Texas",
         "49" = "Utah",
         "50" = "Vermont",
         "51" = "Virginia",
         "53" = "Washington",
         "54" = "West Virginia",
         "55" = "Wisconsin",
         "56" = "Wyoming",
         "66" = "Guam",
         "72" = "Puerto Rico"),
    flushot_yn = recode(flushot_yn,
          "1" = "Yes",
          "2" = "No",
          "7" = "Don't know/not sure",
          "9" = "Refused")
  )
```

BRFSS visual inspection: Where did people get their flu vaccine in 2019? {.storyboard}
=========================================

###  At what kind of place did you get your last flu shot or vaccine?

```{r}
brfss_df %>%
  select(seqno, state, sexvar, genhlth, physhlth, menthlth, hlthpln1, persdoc2, medcost, checkup1,
         flushot7, imfvpla1, urbstat) %>%
  mutate(
    flu_vacc_loc = as.factor(imfvpla1)
  ) %>%
  drop_na(flu_vacc_loc) %>%
  mutate(
    flu_vacc_loc = recode(flu_vacc_loc,
                          '1' = "Doctor's office/HMO",
                          '2' = "Health dept",
                          '3' = "Clinic/health center",
                          '4' = "Senior, rec, or community center",
                          '5' = "Supermarket/pharmacy",
                          '6' = "Hospital",
                          '7' = "ER",
                          '8' = "Workplace",
                          '9' = "Elsewhere",
                          '10' = "Canada/Mexico",
                          '11' = "A school",
                          '77' = "Don't know/not sure",
                          '99' = "Refused"),
  )%>% 
  count(flu_vacc_loc)%>%
  mutate(flu_vacc_loc = fct_reorder(flu_vacc_loc, n)) %>%
    plot_ly(x = ~flu_vacc_loc, y = ~n, color = ~flu_vacc_loc, type = "bar", colors = "plasma") %>%
  layout(
    xaxis = list(title = "Flu vaccine location"),
    yaxis = list(title = "Count (n)")
  )

```

***

We wanted to look at where people were receiving their flu vaccines nationally. By looking purely at counts,
we can see that the leading locations are at doctor's offices and at supermarkets/vaccines, although other
options do exist. Interestingly enough, 3 people reported getting their flu vaccine in either Canada or Mexico.

Another location-related question that came up was if there were any differences in percentages of the surveyed 
population among the regions of the United States. Click along on the other frames in this dashboard to see how 
these percentages compare among the regions.


### Northeastern United States

```{r}

plot_northeast =
 brfss_df %>%
  filter(
    state == c("Connecticut","District of Columbia", "Delaware","Maine", "Massachusetts", "Maryland",
               "New Hampshire", "New Jersey", "New York", "Pennsylvania",
               "Rhode Island", "Vermont")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "Refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="stack", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "Proportion response",
    title = "Frequency of flu vaccinations in the Northeastern region of the U.S."
  )+
  scale_fill_discrete(name = "Flu shot?")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) 
ggplotly(plot_northeast)

```

***
Looking through the states in the region, what trends do you see?

- The split seems to be approximately equal between those who did/did not get the flu shot
- How do these trends compare with the other regions?
- Toggle over the graph to get the exact number of the proportion response
- Of note: this graphic excludes missing values, ie people that did not answer or refused we not included

### Midwestern United States

```{r}

#Midwest
plot_midwest =
  brfss_df %>%
  filter(
    state == c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota",
               "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota",
               "Wisconsin")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "Refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="stack", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "% Response",
    title = "Frequency of flu vaccinations in the Midwestern region of the U.S"
  )+
  scale_fill_discrete(name = "Flu shot?")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
 
ggplotly(plot_midwest) 

```   

***
Looking through the states in the region, what trends do you see?

- The split seems to be approximately equal between those who did/did not get the flu shot
- How do these trends compare with the other regions?
- Toggle over the graph to get the exact number of the proportion response
- Of note: this graphic excludes missing values, ie people that did not answer or refused we not included

### Western United States

```{r}
plot_west =
 brfss_df %>%
  filter(
    state == c("Alaska", "Arizona", "California", "Colorado","Hawaii", "Idaho", "Montana",
               "Nevada", "New Mexico", "Oregon", "Utah", "Washington",
               "Wyoming")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "Refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="stack", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "Proportion response",
    title = "Frequency of flu vaccinations in the Western region of the U.S."
  )+
  scale_fill_discrete(name = "Flu shot?")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
ggplotly(plot_west)

```

***
Looking through the states in the region, what trends do you see?

- The split seems to be approximately equal between those who did/did not get the flu shot
- How do these trends compare with the other regions?
- Toggle over the graph to get the exact number of the proportion response
- Of note: this graphic excludes missing values, ie people that did not answer or refused we not included

### Southern United States

```{r}
plot_south =
 brfss_df %>%
  filter(
    state == c("Alabama","Arkansas", "Florida", "Kentucky", "Louisiana", 
               "Mississippi", "North Carolina", "Oklahoma", "South Carolina",
               "Tennessee", "Texas", "Virginia", "West Virginia")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "Refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="stack", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "Proportion response",
    title = "Frequency of flu vaccinations in the Southern region of the U.S."
  )+
  scale_fill_discrete(name = "Flu shot?")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))

ggplotly(plot_south)
```

***
Looking through the states in the region, what trends do you see?

- The split seems to be approximately equal between those who did/did not get the flu shot
- How do these trends compare with the other regions?
- Toggle over the graph to get the exact number of the proportion response
- Of note: this graphic excludes missing values, ie people that did not answer or refused we not included


**Using the BRFSS dataset, we visually inspected only a few variables that could influence decision to vaccinate. Generally,
it is important to visually inspect and explore the data prior to performing statistical analyses. The questions that this graph 
and that the other plots in this dashboard beg the questions:**

- Is there a significant difference between getting and not getting a flu shot depending on your geographic region?
- Is there a significant difference between getting and not getting a flu shot depending on location accessibility?
- What other variables might be influential in an individuals decision to get vaccinated?

And finally, we would also be interested in what *motivated* people, ie, what were some *qualitative* reasons that drive the data
that we are inspecting visually?


